---
title: "Choose Your Survey Hosting Software"
author: "Rees Morrison"
date: '2022-06-11'
output: pdf_document
draft: no
categories: Surveys
---

```{r setup, include=FALSE, echo=F, warning=F, message=F, cache=F}
knitr:: opts_chunk$set(echo=F, warning=F, message=F, fig.align='center')

library(tidyverse)
library(kableExtra)
library(readxl)
```

If a law firm or law department decides to pursue an online survey, the **project team** will need to select a **software vendor**.  Using the software and hosting services of that vendor, a **team member** or **consultant** will create the questionnaire form that is sent, as a URL link, to invitees.  In this early step of the project, the team might enlist assistance from the information technology department when they evaluate survey software.  So, where do you start to identify potential vendors and how do you choose one?

We will start here with the first inquiry: who markets software to be licensed for for online surveys?  Foraging on the Internet located five blog posts that reviewed the plethora of survey software providers and selected several of them to highlight.

Not all of the posts fully explained how many total hosting vendors they considered or precisely how they decided which vendors met their criterion for inclusion. Zapier's post, at least, states that "we tested dozens of online survey apps."  On the other hand, Zapier also explains that "we review apps that have a Zapier integration," which may exclude otherwise capable offerings.  What they compiled and concluded about the various offerings also ranges widely. 

Table 1 shows the five blog sites, their URL, date of publication, and number of survey vendors included in their review.

```{r reviewSites, eval=FALSE}
sites <- read_excel(path = "C:/Users/rees/Documents/R/Projects/LAWYER Hornbooks/5Surveys/Blogs about SurveysTwo.xlsx")

kable(sites, caption = "Five Blog Posts that Rated Survey Software", align = 'c') %>%   kable_styling(position = "center", full_width = FALSE, latex_options = "hold_position") %>%
  row_spec(0, bold = TRUE, align = "c")

```

The 53 vendors discuss on the five blog posts reduced to a total of 32 offerings reviewed. Only 13 made more than one compilation, of which five of them appeared three or more times.  Table 2 shows that baker’s dozen, sorted by the number of their citations but otherwise listed alphabetically.  A "0" means that the blog post did not include that particular software vendor; a "1" indicates inclusion of that vendor.

```{r ratings, echo=F, warning=F, message=F, cache=F}
ratings <- read_excel(path = "C:/Users/rees/Documents/R/Projects/LAWYER Hornbooks/5Surveys/Blog ratings.xlsx")

# ratings <- ratings[1:13, c(1, 3:8)]

ratings[is.na(ratings)] <- 0

colnames(ratings) <- c("Vendor", "Hubspot", "SaaSlist", "Zapier", "PCMag", "TechRadar", "Cites")

kable(ratings, caption = "Five Ratings of Survey Software (2 or more cites)", align = c(rep('c', 7))) %>% 
  column_spec(7, background = "#90ee90 ") %>% 
  kable_styling(position = "center", full_width = FALSE, latex_options = "hold_position") %>%
  row_spec(0, bold = TRUE, align = "c")

```

Table 3 shows all the remaining vendors that were only included in a single blog-post list.

<!-- ```{r ratings2, echo=F, warning=F, message=F, cache=F, eval=FALSE} -->
<!-- ratings2 <- read_excel(path = "C:/Users/rees/Documents/R/Projects/LAWYER Hornbooks/5Surveys/Blogs about SurveysTwo.xlsx") -->

<!-- ratings2 <- ratings2[14:32, c(1, 3:8)] -->

<!-- ratings2[is.na(ratings2)] <- 0 -->

<!-- colnames(ratings2) <- c("Vendor", "Hubspot", "SaaSlist", "Zapier", "PCMag", "TechRadar", "Cites") -->

<!-- kable(ratings2, caption = "Five Ratings of Survey Software (1 cite)", align = c(rep('c', 7))) %>%  -->
<!--   column_spec(7, background = "#90ee90 ") %>%  -->
<!--   kable_styling(position = "center", full_width = FALSE, latex_options = "hold_position") %>% -->
<!--   row_spec(0, bold = TRUE, align = "c") -->

<!-- ``` -->

These posts can serve as a starting point for a vendor selection process.  I won't venture into trying to evaluate the dozens of survey software options now available on the market.  I will, however, end with two observations from the posts.  Legal-related surveyors probably won't want Surveybot for Facebook surveys, YesInsights with its single question format, or SurveySparrow's chatbot surveys.  I also noted that SaaS makes clear that the companies it includes pays them for inclusion.